---
title: "R Ladies Geospatial Tutorial Extras"
output: html_document
---


```{r}
library(scales) ## commas in legend
library(extrafont) ##fonts for map
loadfonts(device = "win") ### fonts for map
library(ggsn) ##scalebar
library(rgdal) ###used for projections and importing data
library(ggplot2) ##creating the map
library(sf) ## data is stored as sf dataframes
library(data.table)
library(dplyr) ##data transformations
library(stringr)
```

```{r}
#### destination for all unzipped folders
zip_dl = "C:/Users/argun/Documents/Shapefiles/NYC/Testing"


##create a temporary folder to unzip (only do this once, not for every file)
temp_folder= paste0(zip_dl, "/temp")


###final folder for all shapefiles
###create new folder (only do this once, not for every file)
new_dir <-  paste0(zip_dl, "/shps")

```

### Working With Files from a Geodatabase

### We're Going to Work with LION, which is a representation of NYC's Street Network and Other Boundaries

### Reading in Files from a Geodatabase

A geodatabase contains multiple files, so you need to know which file you're reading in. In this case, we only want NYC's road network

```{r}

##read lion from gdb
download.file("https://www1.nyc.gov/assets/planning/download/zip/data-maps/open-data/nyclion_21d.zip", 
              destfile = paste0(zip_dl, "/nyclion_21d.zip"),
              mode = "wb")

unzip(paste0(zip_dl, "/nyclion_21d.zip"),
      exdir=new_dir)


gdb <-  "C:/Users/argun/Documents/Shapefiles/NYC/Testing/shps/lion/lion.gdb" ## make path an object

### let's check out what's in the file geodatabase
ogrListLayers(gdb) ### create an object that is a list of layers

```

### Based on the metadata, we know LION is the road network

```{r}

lion <- st_read(dsn=gdb,layer="lion") ## read in data
lion_cast <- st_cast(lion, "MULTILINESTRING")
lion_condensed <- lion_cast %>%
  ### remove railroads, water edge, census block boundary, paper street
  ### district boundary, alley, and ferry route
  filter(!FeatureTyp %in% c("1", "2", "3", "5", "7", "8", "A", "F" )) %>%
  filter(!RB_Layer %in% c("R", "S", "F")) ## remove roadbed, suppressed, and fake segments
lion_condensed <- st_transform(lion_condensed, 2263)
```
### Checking LION Geometry before st_cast command above
```{r}
st_geometry_type(lion, by_geometry = FALSE)
```
### LION Geometry after st_cast
### If you want to perform spatial operations on a sf file, the geometry type needs to be consistent
```{r}
st_geometry_type(lion_cast, by_geometry = FALSE)
```
```{r}
rm(lion)
rm(lion_cast)
```

### Read In All Other Data - Process for Downloading boroughs, parks, and ntas is in the main file
### Here I also added parks properties

```{r}
boroughs <- st_transform(st_read(new_dir, "boroughs" ), 2263)
ntas <- st_transform(st_read(new_dir, "ntas"), 2263)


### restaurant data
rest <- fread("C:/Users/argun/Documents/Shapefiles/NYC/Testing/open_restaurants.csv", header = TRUE)

###remove rows without lat/long data
rest <-  rest[!is.na(rest$Latitude) | !is.na(rest$Longitude), ]

###make the imported table in to a sf dataframe
rest_sf <- st_as_sf(rest , coords = c("Longitude", "Latitude"))


###set geographic coordinate system and projected coordinate system
sf_proj <- st_transform(st_set_crs(rest_sf, value = 4326), "EPSG: 2263")

##PCS
#sf_proj <- st_transform(sf_proj, "EPSG: 2263")

###remove duplicates
rests_sf <- unique(sf_proj, by = "objectid") ## sf version for spatial join


### Parks Properties - Download and import

####NTAS, eventually derive parks from this
download.file("https://data.cityofnewyork.us/api/geospatial/enfh-gkve?method=export&format=Shapefile", 
              destfile = paste0(zip_dl, "/parks_properties.zip"),
              mode = "wb")

unzip(paste0(zip_dl, "/parks_properties.zip"),
      exdir=temp_folder)

old_files <- list.files(temp_folder, pattern = "geo*", full.names = TRUE)

new_names = gsub("^.*?\\.","parks_properties.",old_files)

file.copy(from=old_files, to=paste0(new_dir, "/", new_names))
file.remove(old_files)
list.files(new_dir)


parks_properties <- st_transform(st_read(new_dir, "parks_properties"), 2263)


###remove excess sf dataframes
rm(rest)
rm(rest_sf)
rm(sf_proj)
gc()
```


### Let's Make a Closeup Map. I picked Crown Heights North
### In order to get the proper boundaries, you need the bounding box of the neighborhood; the bounding box defines the coordinates of the neighborhood

```{r}

bounding_box = st_bbox(ntas[ntas$ntaname=="Crown Heights North",])

####data for zoomed in map
ntas_cropped <- st_crop(ntas, bounding_box)
rest_cropped <- st_intersection(rests_sf, ntas_cropped)
streets_cropped <- st_intersection(lion_condensed, ntas_cropped)
parks_properties_cropped <- st_intersection(parks_properties, ntas_cropped) %>%
  ###remove excess parks area like the green space on Eastern Parkway
  subset(!typecatego %in% c("Mall", "Parkway"))
```

### What does this look like on a map?

```{r}
ggplot()+
  geom_sf(data=ntas_cropped, fill="white")+
  geom_sf(data=parks_properties_cropped, fill="#a1c690", color="lightgrey")+
  geom_sf(data = streets_cropped, color="#f2f2f2", size=.005)+
  geom_sf(data=rest_cropped, color="red", size=.2)+
  theme(panel.background = element_rect(fill = "white"),
        axis.ticks = element_blank(),
        axis.text = element_blank(),
        panel.grid = element_line(color = "white", size = 0.8))
```




### Watch What Happens When you Try and Add Street Labels...chaos!

### This is because each street is not one long line in lion. Each segment between two intersections is a record in the data, so when every segment is labeled, chaos ensues.

```{r}
ggplot()+
  geom_sf(data=ntas_cropped, fill="white")+
  geom_sf(data=parks_properties_cropped, fill="#a1c690", color="lightgrey")+
  geom_sf(data = streets_cropped, color="#f2f2f2", size=.005)+
  geom_sf_text(data=streets_cropped, 
               aes(label=Street, family="serif"), colour = "black", size=2)+
  geom_sf(data=rest_cropped, color="red", size=.2)+
  theme(panel.background = element_rect(fill = "white"),
        axis.ticks = element_blank(),
        axis.text = element_blank(),
        panel.grid = element_line(color = "white", size = 0.8))
```

### Processing for Labeling Roads

###Each section of the road between two intersections is a separate line in the data. Therefore, there are several lines that represent each street

### The spatial function st_union() allows us to merge these lines based on a column. Now each street name is represented by a single row

```{r}
###union roads for the purpose of labeling
roads.out <- streets_cropped %>% 
  group_by(Street) %>% 
  summarize(geometry = st_union(SHAPE))
```

### In order to make sure the label appears in the right place on the map, you need to manually adjust the the x/y coordinates of the label using geom_sf_text(). This is a very tedious process; the main drawback of using R as a geospatial tool is that labeling is time-consuming. It took me several tries to find a place where the label was sitting close enough to the street without overlapping any points.

```{r}
ggplot() + 
  geom_sf(data=streets_cropped, aes(color="streets"), show.legend = "line", size=.005)+
  geom_sf(data=parks_properties_cropped, fill="#a1c690", color=NA)+
  geom_sf_text(data=roads.out[roads.out$Street=="ATLANTIC AVENUE",], 
               aes(label=Street, family="serif"), colour = "black", size=2, angle=-5
               ,nudge_x=1600, nudge_y=-35)+
  geom_sf_text(data=roads.out[roads.out$Street=="FRANKLIN AVENUE",], 
               aes(label=Street, family="serif"), colour = "black", size=2, 
               angle=80 ,
               nudge_x = -70, nudge_y = 350)+
  geom_sf_text(data=roads.out[roads.out$Street=="EASTERN PARKWAY",], 
               aes(label=Street, family="serif"), colour = "black", size=2, angle=-5,
              nudge_x = 5000, nudge_y=-750)+
  geom_sf(data=rest_cropped, aes(color="Restaurant Permits"), show.legend = "point", size=1.2)+
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank(),
        panel.background = element_blank(),
        plot.background = element_rect(fill = "white"),
        plot.title= element_text(colour="black", size = 20, face = "bold", hjust=0.5,),
        legend.position = "right") +
  theme(legend.title = element_blank(),
        legend.spacing.y = unit(0, "mm"), 
        panel.border = element_rect(colour = "black", fill=NA),
        aspect.ratio = 1, axis.text = element_text(colour = 1, size = 12),
        legend.background = element_blank(),
        legend.box.background = element_rect(colour = "black"))+
  labs(title = "Open Restaurant Permits \n Crown Heights North")+
  xlab("") + 
  ylab("")+
  scale_colour_manual(values = c("Restaurant Permits" = "red", "streets"="darkgrey"),
                      guide = guide_legend(override.aes = list(linetype =  c("blank", "solid"),
                                                              shape=c(16, NA))))+
  scalebar(ntas_cropped, dist = .5, dist_unit = "mi", location="bottomleft",
             transform = FALSE, model = "WGS84")
  
```



